---
title: "BST260WHR project"
author: "Sung-An Tai, Ellen Kao, Flora Chang, Allina Chang"
date: "2020/12/9"
output: html_document
---
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl) 
library(gsheet)
library(MASS)
library(ggplot2)
library(grid)
library(gridExtra)
library(asbio)
library(rvest)
library(maps)
library(randomForest)
library(caret)
library(sf)
library(ggrepel)
library(ggspatial)
library(gsheet)
```

# Overview and Motivation
In our own ways, we are all seeking happiness, not only for us, but also for our loved ones. We wanted to choose a topic that is relatable for everyone. Thus, we looked into the World Happiness Report. The World Happiness Report is a landmark survey of the state of global happiness that ranks countries by how happy their citizens perceive themselves to be. Especially during this new COVID-19 reality that we are living in, external factors can often shake our happiness. We thought it would be interesting to dig deeper into the predictors of happiness. With team members located in Taiwan and the US, we also wanted to choose a topic that compares and contrasts across countries. Through this project, we hoped to gain insight about how happiness trends vary across different countries, how different variables may or may not explain happiness, and how happiness can be predicted in the future. 

# Related Work
The world map on the midterm inspired us to visualize data across the entire world, and current day events revolving COVID-19 led us to choose a topic that is largely impacted by the consequences of the pandemic. 

# Initial Questions
Our initial question was: How does happiness vary across the world? After beginning to look into the data and starting to visualize the results, additional questions arose such as:
Does geographical location play a role in happiness levels?
Are countries that are closer to the equator happier?
Which variables predict happiness? 
How does happiness change as a function of time?
Are first-world countries happier than second- or third-world countries?

# Data Visualization

First, we read in the data pulled from the 2020 World Happiness Report. We decided to use the datasets for Table 2.1 and Figure 2.1 in the report. 
```{r}

#Figure 2.1
url20 <- "https://docs.google.com/spreadsheets/d/1vkHkvnZFhtekjkYbF-y7uVo4oE36ROaqJx4lxVIkdq4/edit?usp=sharing"
dat20 <- gsheet2tbl(url20, sheetid = NULL)

#Table 2.1
dat2_1 <- "https://docs.google.com/spreadsheets/d/1bAzkkXU3W7LALAzP2cnaahbf-s52kr8GvjJ3pIaK9Cs/edit?usp=sharing"
whr2_1 <- gsheet2tbl(dat2_1, sheetid = NULL)
```

We chose to visualize happiness levels across the world in 2019 and make some conclusions about whether or not happiness is regionally correlated. We also thought it would be interesting to visualize how discrepant happiness levels could be within continents. 

We also visualized happiness across the years as well as the relation of the Ladder Score to other variables in a [Shiny app](https://changallina.shinyapps.io/bst260finalproject/) format.
Our Shiny app has two tabs, the first visualizes happiness trends over the years and the second visualizes how happiness correlates with other country characteristics.

1. Happiness trends: The user is allowed to select up to six countries at once. Although not every country has Country Ladder score data for every year, we chose to make the x-axis flexible to input and display the entire range of years (2005 to 2019) when appropriate. An interesting conclusion is that the United States has actually been on a slight downward trend in terms of overall "happiness." Some [reports](https://www.washingtonpost.com/nation/2019/03/21/americans-are-unhappiest-theyve-ever-been-un-report-finds-an-epidemic-addictions-could-be-blame/) attribute this decline to an "epidemic of addiction," which covers everything from substance abuse to social media. In turn, these addictions are causing higher rates of depression and general unhappiness among the population. Our Shiny app makes it easy to visualize, compare, and contrast happiness in different countries and across multiple years. These trends can be used as starting points of future research.

2. Dynamics of variables: In the second tab, the user is allowed to select up to three countries at once as well as any continuous range of years between 2005 and 2019. This tab aimed to visualize the relationship between happiness and other variables, including GDP, healthy life expectancy at birth, social support, freedom to make life choices, generosity, and perceptions of corruption. The y-axis represents the value of the Cantril Ladder score while the x-axis represents the value of the selected variable. Based on this plot, users can visualize the correlation between a specified variable and happiness while also viewing how these trends may differ across countries. This visualization provides insight into the strength and role with which a variable may play in relation to overall happiness. For example, when "Italy", "Taiwan", and the "United States" are selected along with GDP, it's apparent that GDP is quite positively correlated with happiness scores. This is consistent with [research](https://psycnet.apa.org/record/2009-09430-011) indicating that judgment measures, such as the Cantril Ladder score, are strongly associated with income as well as long-term changes of national income. The higher GDP per capita, the higher the Ladder score. Of course, we know that GDP is not the only predictor of happiness (or even close to!), so we still need to consider other possible predictors of happiness. 

Our next visualization involved creating and filling in a world map according to 2019 Ladder scores. In this first code chunk, we combined our dataframe containing data from Figure 2.1 of the 2020 World Happiness Report with a world map dataframe. In order to do so, we had to rename some of the countries so that they would match up between the two sets of data. Additionally, we wanted to include Hong Kong as its own country/region instead of treating it as a subregion within China, as the world map dataset did. 

To make our final world map easier to interpret, we incremented Cantril Ladder Scores by 0.5. Since there were no ladder scores below 2, that was the lower bound of our first category. 
```{r}
world_map = map_data("world")

dat20 <- dat20 %>%
  mutate(new_names = recode(`Country name`,
    "United Kingdom" = "UK",
    "United States" = "USA",
    "Taiwan Province of China" = "Taiwan",
    "Trinidad and Tobago" = "Trinidad",
    "North Cyprus" = "Cyprus",
    "Hong Kong S.A.R. of China" = "Hong Kong",
    "Palestinian Territories" = "Palestine",
    "Congo (Brazzaville)" = "Democratic Republic of the Congo"
  )
  )

world_map <- within(world_map, region[region == "China" & subregion == "Hong Kong"] <- "Hong Kong")

world_map <- world_map %>%
  filter(region != "Antarctica") %>% 
  rename("Country" = region)

dat20 <- dat20 %>%
  rename("Country" = new_names)

combined_countries <- full_join(dat20, world_map)

combined_countries <- combined_countries %>% mutate(manual_fill = cut(`Ladder score`, 
                         breaks = c(2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.5, 8, Inf),
                         labels = c("(2, 2.5]", "(2.5, 3]", "(3, 3.5]", "(3.5, 4]", "(4, 4.5]",
                                    "(4.5, 5]", "(5, 5.5]", "(5.5, 6]", "(6, 6.5]", 
                                    "(6.5, 7]", "(7, 7.5]", "(7.5, 8]", ">8"), 
                         right = TRUE))

whr2_1 <- whr2_1 %>% rename(GDP = `Log GDP per capita`,
                            "LifeExpectancy" = `Healthy life expectancy at birth`,
                            "SocialSupport" = `Social support`,
                            "PerceptionsOfCorruption" = `Perceptions of corruption`)
```

In this section of code, we identified the lowest and highest Cantril ladder scores within each continent. As Australia can be considered as its own continent, we did not indicate the lowest-ladder score in that region.
```{r}
topcities <- data.frame(Country = c("Taiwan", "Australia", "Mauritius", "Chile", "Canada", "Finland"), 
    lat = c(23.6978, -43.23242, -20.48486, -33.447487, 56.1394, 60.04068), 
    lng = c(120.9605, 147.3488, 57.65127, -70.673676, -106.3468, 20.61133),
    level = c("happiest", "happiest", "happiest", "happiest", "happiest", "happiest"))

(topcities <- st_as_sf(topcities, coords = c("lng", "lat"), remove = FALSE, 
    crs = 4326, agr = "constant"))

lowcities <- data.frame(Country = c("Venezuela", "Haiti", "South Sudan", "Ukraine", "Afghanistan"), 
    lat = c(6.4238, 18.9715, 6.8770, 48.3794, 33.9391), 
    lng = c(-66.5897, -72.2852, 31.3070, 31.1656, 67.6100),
    level = c("saddest", "saddest", "saddest", "saddest", "saddest"))

(lowcities <- st_as_sf(lowcities, coords = c("lng", "lat"), remove = FALSE, 
    crs = 4326, agr = "constant"))

mostleast <- rbind(topcities, lowcities)
```

We are finally able to build our world map based on Cantril Ladder Scores. Countries of note for 2019 include South Sudan and Finland, two countries that represent the extremes in average Cantrill Ladder score. Happiness is at least somewhat regionally correlated which reflects how neighboring countries may share general characteristics that lead to happiness levels. Most of the countries in Africa and in the Middle East are ranked significantly lower in "happiness" compared to other countries in the world while Europe and the Americas display relatively high ladder scores. 
```{r}
ggplot() +
  geom_polygon(combined_countries, mapping = aes(x = long, y = lat, group = group, fill = manual_fill)) +
  geom_text_repel(data = mostleast, aes(x = lng, y = lat, label = Country, color = level), 
                  size = 3,
                  fontface = "bold",
                  nudge_x = c(25, -1.5, 15, -13, -40, -16, 8, 3, 46, 3, 2),
                  nudge_y = c(-5, -15, 7, 3, -7, 13, 10, 7, -3, 26, 45),
                  min.segment.length = unit(0, "lines")) +
  geom_sf(data = topcities, alpha = 0.5, size = 0.5, color = "salmon3") +
  geom_sf(data = lowcities, alpha = 0.5, size = 0.5, color = "dodgerblue4") +
  ggtitle("Happiness Levels Across the World in 2019 (based on WHR 2020)") +
  scale_fill_manual(name = "Cantrill Ladder Score", 
                     values = c("dodgerblue4",
                                "dodgerblue3", 
                                "lightskyblue2",
                                "paleturquoise3",
                                "lightcyan3",
                                "honeydew2",
                                "khaki1",
                                "gold1",
                                "goldenrod1",
                                "darkgoldenrod2",
                                "salmon3"),
                    na.value = "gray") +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 7)) +
  guides(shape = guide_legend(override.aes = list(size = 5))) +
  theme(panel.grid.major = element_blank(), 
        panel.background = element_blank(),
        axis.title = element_blank(), 
        axis.text = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank())
  
```

# Data Analysis: Regression

```{r, warning=FALSE}
# Read WHR20_DataForFigure2.1.xls (2017~2019 average) and WHR20_DataForTable2.1.xls (2008~2019 with additional variables)
# Main dependent variable: Ladder score, independent variable: `Logged GDP per capita`+`Social support`+`Healthy life expectancy`+`Freedom to make life choices`+`Generosity`+`Perceptions of corruption`
whr20 <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1jVOpMfl94xDmip0NrJEXORoFK4RBfB7leDVzyJnk47w/edit#gid=28299357", sheetid = NULL)
whr_all <- gsheet2tbl("https://docs.google.com/spreadsheets/d/1oMzplKOmjs0hKKmXKqOCM90CvPsS7aCIj1adox0nvNU/edit#gid=1261774380", sheetid = NULL)
```

We first look at the regression model for the pooled data of 2017-2019 
```{r}
# Main model for fig 2.1 in the WHR 2020 report, but they used pooled OLS regression specific for panel analysis, rather than plain multiple regression
mod1 <- lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)
summary(mod1)
```

Then, we tried to see how much variance each variable explained in the data by looking at partial R2.
```{r}
#Logged GPD per capita explained about 5% of the variance in the model. 
lm.with<-lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)

lm.without<-update(lm.with, ~. - `Logged GDP per capita`)

R2_gdp <-  partial.R2(lm.without,lm.with)
R2_gdp

## Social support explained 10% of the variance in the model. 

lm.with<-lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)

lm.without.2<-update(lm.with, ~. - `Social support`)

R2_social_support <-  partial.R2(lm.without.2,lm.with)
R2_social_support


##Healthy life expectancy explained 4.8% of the variance in the model.
lm.with<-lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)

lm.without.3<-update(lm.with, ~. - `Healthy life expectancy`)

R2_healthy_life_expectancy <-  partial.R2(lm.without.3,lm.with)
R2_healthy_life_expectancy

## Freedom to make life choices explained 8% of the variance in the model. 
lm.with<-lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)

lm.without.4<-update(lm.with, ~. - `Freedom to make life choices`)

R2_choice <-  partial.R2(lm.without.4,lm.with)
R2_choice

##Generostiy explained 1% of the variance in the model. 
lm.with<-lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)

lm.without.5<-update(lm.with, ~. - `Generosity`)

R2_generosity <-  partial.R2(lm.without.5,lm.with)
R2_generosity


## Perceptions of corruption explained 2.7% of the variance in the model. 
lm.with<-lm(`Ladder score` ~ `Logged GDP per capita`+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr20)

lm.without.6<-update(lm.with, ~. - `Perceptions of corruption`)

R2_corruption <-  partial.R2(lm.without.6,lm.with)
R2_corruption
```

Based on the partial R squared results, we found the the top 3 variables which explained the most variance in the model were "Social support" (10%), "Freedom to make life choices" (8%), and "Logged GDP per capita" (5%).


After we ran multiple regression  on the data (2017-2019), we found that five out of six variables were significant. The only exception was generosity. Therefore, we have ran some scatter plots to see the relationship between happiness score (ladder score) and each variable. From the scatter plots below, we found that it seems that there was not a linear relationship between  generosity and happiness score and between percpetions of corruption and happiness score.
```{r}
p_1 <- whr20 %>% 
  ggplot(aes(x= `Logged GDP per capita`, y= `Ladder score`)) +
    geom_point() +
    ylab("Happiness score")  +
    xlab("Logged GDP per capita") +
  ggtitle("Ladder vs GDP per Capita")

p_1


p_2 <- whr20 %>% 
  ggplot(aes(x= `Social support`, y= `Ladder score`)) +
    geom_point() +
    ylab("Happiness score")  +
    xlab("Social support") +
  ggtitle("Ladder vs Social Support")

p_2

p_3 <- whr20 %>% 
  ggplot(aes(x= `Healthy life expectancy`, y= `Ladder score`)) +
    geom_point() +
    ylab("Happiness score")  +
    xlab("Healthy life expectancy") +
  ggtitle("Ladder vs Healthy Life Expectancy")

p_3

p_4 <- whr20 %>% 
  ggplot(aes(x= `Freedom to make life choices`, y= `Ladder score`)) +
    geom_point() +
    ylab("Happiness score")  +
    xlab("Freedom to make life choices") +
  ggtitle("Ladder vs Freedom to make life choices")

p_4

p_5 <- whr20 %>%
    ggplot(aes(x= `Generosity`, y= `Ladder score`)) +
    geom_point() +
    ylab("Happiness score")  +
    xlab("Generosity") +
  ggtitle("Ladder vs Generosity")

p_5


p_6 <- whr20 %>% 
  ggplot(aes(x= `Perceptions of corruption`, y= `Ladder score`)) +
    geom_point() +
    ylab("Happiness score")  +
    xlab("Perceptions of corruption") +
  ggtitle("Ladder vs Perceptions of Corruption")
  
p_6


grid.arrange(p_1, p_2, p_3, p_4, p_5, p_6, nrow = 3)
```


We ran some diagnostic plots to check the assumptions for linear regression model(2017-2019 data). Diagnostics seems to fulfill "linearity","normality", and "equal variance" assumptions.
```{r}
# Residuals versus fitted plot
plot(fitted(mod1), residuals(mod1)) 
abline(a=0,b=0,col="pink")

# QQ Norm plot
qqnorm(residuals(mod1))
qqline(residuals(mod1),col="pink")

```


## Multiple regression analyses for 2015 - 2019, separately
```{r}
# Model from 2015 to 2019
modeling <- function(y){
  whr_df <- whr_all %>% filter(year == y) %>% dplyr::select(`Life Ladder`, `Log GDP per capita`,
             `Social support`, `Healthy life expectancy at birth`,
             `Freedom to make life choices`,
             `Generosity`,
             `Perceptions of corruption`) %>% drop_na()
  
  mod <- lm(`Life Ladder` ~ `Log GDP per capita`+
             `Social support`+`Healthy life expectancy at birth`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr_df)
  mod
}
years <- 2015:2019
mods.15to19 <- lapply(years, modeling)
names(mods.15to19) <- paste0("mod.", 2015:2019)
lapply(mods.15to19, summary)
```

We found that generally Logged GDP per capita, Social support, Healthy life expectancy, and Freedom to make life choices are 4 out of the 6 variables that have significant coefficient across 2015-2019, except for social support being insignificant in 2019.


## Model backward selection for 2015 - 2019
```{r}
# backward selection for 2017~2019 model
mod1_bkwd <- step(mod1, direction = "backward")
summary(mod1_bkwd)

# backward selection from 2015 to 2019
backward.select <- function(y){
  whr_df <- whr_all %>% filter(year == y) %>% dplyr::select(`Life Ladder`, `Log GDP per capita`,
             `Social support`, `Healthy life expectancy at birth`,
             `Freedom to make life choices`,
             `Generosity`,
             `Perceptions of corruption`) %>% drop_na()
  
  mod <- lm(`Life Ladder` ~ `Log GDP per capita`+
             `Social support`+`Healthy life expectancy at birth`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr_df)
  mod.backward <- step(mod, direction = "backward")
  mod.backward
}

summary(b2015 <- backward.select(2015))
summary(b2016 <- backward.select(2016))
summary(b2017 <- backward.select(2017))
summary(b2018 <- backward.select(2018))
summary(b2019 <- backward.select(2019))

list(mod.2015 = coef(b2015), mod.2016 = coef(b2016), mod.2017 = coef(b2017), mod.2018 = coef(b2018), mod.2019 = coef(b2019))

```

Backward model selection by year using AIC also showed similar conclusion that Logged GDP per capita, Social support, Healthy life expectancy, and Freedom to make life choices are the more influential variables to Cantril Ladder Score.

However, we believe that GDP per capita is potentially related to the other variables, including social support levels, overall life expectancy, freedom to make life choices, generosity, and perception of corruption. This can be guessed by looking at pairwise distribution of each variable:
```{r}
whr20_few <- whr20 %>% dplyr::select(`Ladder score`, `Logged GDP per capita`,
             `Social support`,`Healthy life expectancy`,
             `Freedom to make life choices`,
             `Perceptions of corruption`,
               Generosity) %>% drop_na()
pairs(whr20_few[-1])
```

We find that GDP per capita may be highly correlated with social support, life expectancy, and freedom to make life choices. It may be a good idea to stratify by GPD per capital to further explore if and how the relationship between these variables and the Cantril Ladder score changes.


## Stratifying by GDP per Capita
```{r}
# Categorize by GDP, divide into 4 quadrants (gdp_cat = 0,1,2, or 3)
whr_gdp <- whr_all %>% filter(!is.na(`Log GDP per capita`))
q_gdp <- whr_gdp %>% summarise(q1 = quantile(`Log GDP per capita`)[2], q3 = quantile(`Log GDP per capita`)[4], mean = mean(`Log GDP per capita`))
quantile(whr_gdp$`Log GDP per capita`)
whr_gdp <- whr_gdp %>% mutate(gdp_cat = cut(`Log GDP per capita`, breaks = c(-Inf, q_gdp[1], q_gdp[3], q_gdp[2], Inf), labels=c(0,1,2,3)))

# Plot gdp-stratified Cantril ladder score v.s. Generosity, since it was the least significant predictor from previous models
whr_gdp %>% filter(gdp_cat == 3, year %in% 2015:2019) %>% 
  ggplot(aes(x = `Generosity`, y = `Life Ladder`)) + 
  geom_point(color = "blue")+
  facet_wrap(.~year)+
  ggtitle("Cantril Ladder vs Generosity at highest GDP category")

whr_gdp %>% filter(gdp_cat == 2, year %in% 2015:2019) %>% 
  ggplot(aes(x = `Generosity`, y = `Life Ladder`)) + 
  geom_point(color = "green")+
  facet_wrap(.~year)+
  ggtitle("Cantril Ladder vs Generosity at 2nd GDP category")

whr_gdp %>% filter(gdp_cat == 1, year %in% 2015:2019) %>% 
  ggplot(aes(x = `Generosity`, y = `Life Ladder`)) + 
  geom_point(color = "red")+
  facet_wrap(.~year)+
  ggtitle("Cantril Ladder vs Generosity at 3rd GDP category")

whr_gdp %>% filter(gdp_cat == 0, year %in% 2015:2019) %>% 
  ggplot(aes(x = `Generosity`, y = `Life Ladder`)) + 
  geom_point(color = "purple")+
  facet_wrap(.~year)+
  ggtitle("Cantril Ladder vs Generosity at 4th GDP category")
```

We found that across 2015-2019, Generosity seems to be possibly correlated with Cantril Ladder score among the countries with the highest quadrant of GDP per capital.

Lets further create models from the data restricted to the highest quadrant of GDP per capita:
```{r}
# Create model from highest GDP per capita category stratified data
model_gdp3_year <- function(y){
  whr_df <- whr_gdp %>% filter(gdp_cat==3, year == y) %>% dplyr::select(`Life Ladder`, `Log GDP per capita`,
             `Social support`, `Healthy life expectancy at birth`,
             `Freedom to make life choices`,
             `Generosity`,
             `Perceptions of corruption`) %>% drop_na()
  
  mod <- lm(`Life Ladder` ~ `Log GDP per capita`+
             `Social support`+`Healthy life expectancy at birth`+
             `Freedom to make life choices`+
             `Generosity`+
             `Perceptions of corruption`,
           data = whr_df)
  mod
}

models_gdp3<- lapply(years, model_gdp3_year)
names(models_gdp3) <- paste0("models_gdp3.", 2015:2019)
lapply(models_gdp3, summary)
```

There are 2 findings:
1. In some years, the coefficients of either social support, life expectancy, or freedom to make life choices that were believed to be correlated with GDP per capita became insignificant due to stratification on GDP per capita and their co-linearity.
2. The coefficient of generosity was originally insignificant in the models across 2015 - 2019. After restricting to the stratum of the highest GDP per capita, it became significant in the models representing years 2015 - 2018. This may indicate that among the higher income countries in the world, generosity level may have significant influence on happiness level, whereas in lower income countries, generosity may simply be a minor factor affecting happiness.

We also want to know how would the model behave after changing GDP per capita into categorized variable. We'll apply it to the pooled data in 2017-2019.

## Change GDP per capita into categorized variable
```{r}
# first create gdp category in whr20 data
whr20_gdp <- whr20 %>% filter(!is.na(`Logged GDP per capita`))
q20_gdp <- whr20_gdp %>% summarise(q1 = quantile(`Logged GDP per capita`)[2], q3 = quantile(`Logged GDP per capita`)[4], mean = mean(`Logged GDP per capita`))
whr20_gdp <- whr20_gdp %>% mutate(gdp_cat = cut(`Logged GDP per capita`, breaks = c(-Inf, q20_gdp[1], q20_gdp[3], q20_gdp[2], Inf), labels=c(0,1,2,3)))

# modeling
mod1_gdp_gen <- lm(`Ladder score` ~ gdp_cat+
             `Social support`+`Healthy life expectancy`+
             `Freedom to make life choices`+
             `Perceptions of corruption`+
               Generosity,
           data = whr20_gdp)
summary(mod1_gdp_gen)
```

The result of the model showed that only the highest GDP per capita has a significant coefficient compared to the lowest GDP per capita. Category 1 and 2 showed no difference of their effect on Cantril Ladder score compared to the lowest category. The coefficients of social support, life expectancy, and freedom to make life choices remained significant as the model with continuous GDP per capita variable. Note that according to R-squared value, this model (adj-R2 = 0.7555) performs better than the original model(adj-R2 = 0.738).

# Data Analysis: Random Forest

Here, we used a Random Forest approach to predict Life Ladder scores based on other variables present in the World Happiness Report dataset. We chose to use a subset of the original variables (GDP, social support, life expectancy, freedom to make life choices, generosity, perceptions of corruption, positive affect, negative affect, and confidence in government) due to the large amount of missing values in the remaining variables such as GINI index. However, there were some variables such as "Freedom to make life choices" and "Perception of corruption" that had missing values every year for a few countries such as China. We hypothesized that this was due to lack of reporting from the country, which itself has some implications. Therefore, we set missing values for those specific variables to 0. There remained a few missing values scattered across the other variables which were dropped from the data frame. 
We chose to use an 80/20 split for training the data and decided this was an appropriate balance between having more data to train on than a simple 50/50 while leaving enough data in the test_set so that it is still representative of the population.
```{r}
machine_whr <- whr2_1 %>% dplyr::select(3:12)

machine_whr$`Confidence in national government`[is.na(machine_whr$`Confidence in national government`)] <- 0
machine_whr$PerceptionsOfCorruption[is.na(machine_whr$PerceptionsOfCorruption)] <- 0

full_machine_whr <- machine_whr %>%
  drop_na() %>%
  rename(FreedomLifeChoices = `Freedom to make life choices`,
         PositiveAffect = `Positive affect`,
         NegativeAffect = `Negative affect`,
         Confidence = `Confidence in national government`,
         LifeLadder = `Life Ladder`)

set.seed(1)

train_set <- sample(1:nrow(full_machine_whr), as.integer(nrow(full_machine_whr)*0.8))

fit_forest <- randomForest(formula = LifeLadder ~., data = full_machine_whr, subset = train_set)

predict_forest <- predict(fit_forest, newdata = full_machine_whr[-train_set, ])

test = full_machine_whr[-train_set, "LifeLadder"]

plot(predict_forest, pull(test, 1))

mean((predict_forest - pull(test, 1))^2)

```

The MSE on the test set for this random forest is 0.1848882 which seems low relative to the Life Ladder scores. This is a suggestion that the model is fairly good at predicting life ladder scores from the information we gave it (GDP, social support, life expectancy, freedom to make life choices, generosity, perceptions of corruption, positive affect, negative affect, and confidence in government). 
The figure plots predicted values by the model against the true values from the original dataset. The points are centered around a slope of 1 (absolute error of 0), which shows that the model is not consistently under-predicting or over-predicting life ladder scores given the other variables listed above. 
This model may be useful in countries that are missing survey data from the Gallup World Poll that has Cantril Ladder scores, but still want to estimate the average "happiness" levels of their population, given the appropriate variables. 

